---
title: "Tiles, images, and basemaps"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5,
  dpi = 150
)
```

There are two ways to get a picture of a map out of ArcGIS. A **cached**
service hands you square tiles it drew in advance, and a **dynamic** service
draws the exact extent you ask for.

`arcgistiles` does both, and `as_rast()` turns either into a georeferenced
`SpatRaster`. That is the point of the package: a basemap you can plot your own
data on top of.

```{r setup}
library(arcgistiles)
library(terra)
library(sf)
```

## Opening a service

`map_server()` reads a service and its tiling scheme.

```{r}
ms <- map_server(world_topo_map_url())
ms
```

`Levels: 0-23` means the cache holds 24 zoom levels. `lods()` shows what each
one is worth in map units.

```{r}
head(lods(ms), 5)
```

## Picking a level

Level numbers are not a unit anyone thinks in. Give `get_tiles()` a `size` in
pixels instead and it picks the level that renders your extent closest to it.

```{r}
nc <- read_sf(system.file("shape/nc.shp", package = "sf"))

level_for_size(ms, st_bbox(nc), c(1600, 800))
```

## Tiles for an extent

`tile_grid()` is arithmetic on the tiling scheme, so you can see the cost of a
request before making it. Nine tiles here, one row per tile.

```{r}
tile_grid(ms, st_bbox(nc), level = 7L)
```

`get_tiles()` downloads them in parallel and `as_rast()` merges them into one
raster in the service's CRS.

```{r}
tiles <- get_tiles(ms, st_bbox(nc), size = c(1600, 800), progress = FALSE)
tiles

basemap <- as_rast(tiles)
```

The counties land on the coastline, which is the only real test that the
tiling scheme, the CRS, and the raster extent all agree.

```{r nc-basemap, fig.height = 4}
plotRGB(basemap)
plot(st_transform(st_geometry(nc), st_crs(basemap)), add = TRUE, border = "red")
```

Whole tiles come down, so the raster is always a little wider than what you
asked for. `tileset_bbox()` reports what you actually got.

```{r}
tileset_bbox(tiles)
```

## Dynamic services

A dynamic service draws on request, so you can filter it. `export_map()` takes
layer ids and SQL definition expressions.

```{r}
census <- map_server(census_url())

img <- export_map(
  census,
  c(-104, 35.6, -94.32, 41),
  size = c(600, 400),
  format = "png32",
  layers = 3,
  layer_defs = list(`3` = "POP2000 > 1000000")
)

img
```

Note the extent that came back. The server widened the request to match the
aspect ratio of `size`, which is why `MapImage` carries the extent it drew
rather than the one you sent.

`format = "png32"` above is deliberate. A plain `"png"` arrives as one paletted
band, which `plot()` handles but `plotRGB()` does not.

```{r counties}
plotRGB(as_rast(img))
```

## Basemaps

The basemap styles service catalogue is readable without a token.

```{r}
styles <- basemap_styles("arcgis")

head(styles[c("path", "name", "group")])
```

Fetching a style document itself needs a token with the
`premium:user:basemaps` privilege.

```{r eval=FALSE}
style <- basemap_style("arcgis/navigation", language = "es")
```
