---
title: "Exporting a map series"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 5,
  dpi = 150
)
```

A map series is one map drawn many times, once per page, at a single scale.
ArcGIS Pro calls it a map series or data driven pages.

`map_series()` is the equivalent for a dynamic map service. Give it a set of
extents and it exports one image per extent in parallel.

```{r setup}
library(arcgistiles)
library(sf)

census <- map_server(census_url())
```

## Pages from a grid

`map_grid()` splits one extent into a regular grid. `overlap` adds a shared
margin so adjacent pages can be trimmed and butted together.

```{r}
pages <- map_grid(c(-104, 35.6, -94.32, 41), nrow = 2, ncol = 3, overlap = 0.05)

pages
```

`map_series()` exports them. Every row reports the extent the server actually
drew and the scale it drew it at.

```{r}
series <- map_series(
  census,
  pages,
  size = c(800, 600),
  format = "png32",
  progress = FALSE
)

series
```

All six pages share one scale, which is the whole point. A per page `export_map()`
loop would give you six different scales and an atlas nobody can compare
across.

```{r}
unique(series$scale)
```

## Pages from features

Pages do not have to be a grid. Pass any `sf` object and you get one image per
feature, framed on that feature's bounding box.

```{r}
nc <- read_sf(system.file("shape/nc.shp", package = "sf"))
big <- nc[order(-nc$AREA), ][1:4, ]

county_pages <- map_series(
  census,
  big,
  size = c(600, 400),
  margin = 0.1,
  page_names = big$NAME,
  progress = FALSE
)

county_pages[c("name", "ok", "scale")]
```

Because each county has its own size, these pages do not share a scale. Pass
the features through `map_grid()` on a common extent when scale has to be
fixed.

## Checking the output

`ok` is `FALSE` for a page the service refused, and the rest of the series
still stands. Nothing throws.

```{r first-page}
all(series$ok)

terra::plotRGB(terra::rast(series$path[1]))
```
