---
title: "Vector tiles as sf"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 6,
  dpi = 150
)
```

Vector tiles carry geometry, not pixels. That means you can pull roads, water,
and building footprints out of a basemap and treat them as `sf` data frames.

Tiles are drawn the same way everywhere in the world, so this is a fast route
to reference geometry for anywhere you have no local data for.

```{r setup}
library(arcgistiles)
library(sf)
```

## Downloading tiles

`vector_tile_server()` opens the service and `get_vector_tiles()` downloads
`.pbf` tiles for an extent.

```{r}
vts <- vector_tile_server(open_street_map_url())
vts
```

Vector tile levels are the familiar web map zoom levels. Level 14 is roughly
neighbourhood scale.

```{r}
boston <- wk::rct(-71.12, 42.34, -71.04, 42.38, crs = 4326)

tiles <- get_vector_tiles(vts, boston, level = 14L, progress = FALSE)
tiles
```

## Finding the layers

Layer names come from the service, not from a standard. Ask the tiles what
they hold before naming anything.

```{r}
vector_tile_layers(tiles)
```

## Decoding

`read_vector_tiles()` returns one `sf` data frame per layer, bound across
tiles. Requires the [protolite](https://cran.r-project.org/package=protolite)
package.

```{r}
parts <- read_vector_tiles(tiles, layers = c("road", "water area"), crs = 3857)

parts$road
```

Features arrive clipped to their tile, so a road crossing four tiles arrives as
four features. Dissolve on an attribute if you need whole geometries.

```{r boston-vector-tiles}
plot(st_geometry(parts$`water area`), col = "#a8c4d8", border = NA)
plot(st_geometry(parts$road), col = "grey30", lwd = 0.4, add = TRUE)
```

## Style resources

The service also publishes everything a renderer needs. `vector_tile_style()`
returns a [Mapbox GL style](https://maplibre.org/maplibre-style-spec/)
document with the relative `sprite` and `glyphs` paths resolved to absolute
URLs.

```{r}
style <- vector_tile_style(vts)

names(style)
style$glyphs
```

Fonts and sprites download to disk for use in a renderer outside R.

```{r}
head(vector_tile_fonts(vts))

vector_tile_sprite(vts)
```
